Testing Gravity Using Type Ia Supernovae Discovered by Next Generation Wide Field Imaging Surveys

### 2019-01-13 doc/article.tex
rewrite second half

\documentclass{aastex62}   	% use "amsart" instead of "article" for AMSLaTeX format
\usepackage{graphicx}				% Use pdf, png, jpg, or epsÂ§ with pdflatex; use eps in DVI mode
								% TeX will automatically convert eps --> pdf in pdflatex		
\usepackage{amssymb}
\usepackage{natbib}
\usepackage{grffile}
\usepackage{textcase}
\usepackage{hyperref}
%SetFonts

@@

The precision in  $f\sigma_8$ derived from the baseline WFD ten-year SN~Ia discoveries has been projected by \citet{2017ApJ...847..128H},
from both their RSD and peculiar velocities.
The results are summarized in Table~\ref{tab:howlett}, where a 5\% distance (0.1~mag) uncertainty is assumed for each supernova.
SNe~Ia alone can provide a $2\%$ measurement of $f\sigma_8$ at $z<0.3$ redshifts lower than where galaxy and cluster (through the
kinematic S-Z effect)
RSD measurements are sensitive.  A joint galaxy RSD and SN peculiar velocity will be even more powerful.
SNe~Ia peculiar velocities will measure $f\sigma_8$ at $z<0.3$  as well as galaxy surveys will at $z \sim 0.6$ using RSD. 
This can be compared to the current 15\% uncertainty from 6dF \citep{2017MNRAS.471..839A}, 3\% uncertainty projected
for the combined TAIPAN and WALLABY+WNSHS surveys \citep{2017MNRAS.464.2517H} for low-redshift measurements of $f\sigma_8$.
At larger $z>0.3$ redshifts DESI projects 10\% accuracies.

@@

\section{Designing a Supernova Survey to Measure $\MakeLowercase{f}\sigma_8$}
\subsection{Survey Parameters that Determine Scientific Performance}
\citet{2017ApJ...847..128H} project how well peculiar velocity surveys can measure $f\sigma_8$. 
They perform their analysis in Fourier space, 
where the Fisher information matrix of parameters $\lambda_i$ from random Gaussian field with mean zero and covariance $C(k)$ is
\begin{equation}
F_{ij} = \frac{V}{2}\int \frac{d^3k}{(2\pi)^3} \text{Tr}\left[ C^{-1} \frac{\partial C}{\partial \lambda_i} C^{-1}
\frac{\partial C}{\partial \lambda_j} \right],
\end{equation}
where the covariance for the velocity-velocity correlation
\begin{equation}
C = P_{vv}(k) + \frac{\sigma^2}{n}
\label{cov:eq}
\end{equation}
is dependent on the power spectrum, noise in the velocity measurement dominated by the intrinsic velocity dispersion $\sigma$, and the density of velocity probes
\citep{2017MNRAS.464.2517H}.  
The primary dependence on the growth of structure is through the normalization of the velocity power spectrum, such that 
\begin{equation}
 \frac{\partial P_{vv}}{\partial \lambda} = \text{constant}
\end{equation}
for the parameter choice
$\lambda=(f\sigma_8)^2$.  In the sample-variance limit, $F_{\lambda \lambda} \propto V$ whereas in the shot noise limit $F_{\lambda \lambda} \propto V n^2 \sigma^{-4}$.

@@

\begin{itemize}
\item Solid Angle $\Omega$:
The variance in $f\sigma_8$ is inversely proportional to solid angle $\Omega$.  The baseline
LSST survey covers 18,000 sq.~deg, at $-75 \lesssim \delta \lesssim15$ avoiding the Galactic plane.  Larger solid-angle coverage for $z<0.3$ SN~Ia searches, beyond
the LSST baseline, benefits peculiar-velocity science.  Complementary northern-hemisphere surveys and LSST-expanded or independent
coverage of the southern equatorial pole, could significantly increase
the sky coverage with a corresponding decrease in the variance of $f\sigma_8$.
\item Redshift Depth $z_{max}$:
For the number densities generated by LSST and the SN~Ia intrinsic magnitude dispersion, the noise is not sample-variance
limited.  We have run simulations of LSST SN surveys
and calculated the signal-to-noise (STON) of $f\sigma_8$ from the peculiar-velocity correlations. These simulations show, as seen in the left plot
of  Figure~\ref{notshot:fig}, that an increase in the number density of tracers leads to a significant improvement in STON.
It is also worth noting that sample variance is not negligible, as seen by the non-linear dependence of the  STON on $\sigma_M^2/n$.

\begin{figure}
\centering
\includegraphics[width=0.48\textwidth]{../outcosmo/notshot.png}
\includegraphics[width=0.48\textwidth]{../outcosmo/perSN.png}
\caption{Left: STON of $f\sigma_8$ from simulated LSST SN peculiar-velocities only (no RSD).  There is a small level of scatter in the plots idue to the random
realizations made in simulating the surveys.
For several limiting redshift depths, the STON is plotted as a
function of $\sigma_M^2/n$, normalized to intrinsic magnitude dispersion $\sigma_M=0.1$~mag and 10-year LSST number densities.
Right: The effective contribution to the STON per supernova, quantified as STON divided by the root of the number of tracers.
\label{notshot:fig}}
\end{figure}

@@

@@

%
%
%\appendix
%\section{Reanalysis in Configuration Space}
%
%This Appendix describes a new analysis used to project precisions on $f\sigma_8$ using peculiar velocities derived from SNe~Ia.
%This analysis may or may not be useful for the Decadal Survey White Paper, but will be helpful for LSST planning.
%While there have been a number of articles on the subject,
%our analysis brings a higher level of fidelity than sought previously.  We simulate SNe~Ia hosted by galaxies in a mock galaxy
%catalog. The numbers of SNe are sufficiently small to allow fast evaluations of the likelihood, which enable the determination of parameter
%posteriors using MCMC on reasonable computing timescales.   We can use our machinery to 
%compare different survey parameters, such as redshift depth, total numbers of supernovae, non-uniform spatial density,
%solid angle/survey geometry, and SN~Ia intrinsic magnitude dispersion.
%Adding the cross-correlation between galaxy-count and peculiar-velocity surveys is active work, so we can
%quantify the suppression of sample variance achieved when considering matter-densities and velocities 
%within the same volume \citep{2007PhRvL..99h1301G}.
%
%\section{Simulated Data}
%The cosmoDC2 (v1.0) covering 706 sq.~deg is used\footnote{We previously used
%Buzzard (v1.6) galaxy catalog is used, but it was found to give peculiar velocities inconsistent with general relativity.  As we will see,
%cosmoDC2 (v1.0) is consistent with GR. }. 
%The catalog is based on a Flat $\Lambda$CDM model with $H_0=71$~km~s$^{-1}$,  $\Omega_M=0.265$, $\Omega_B=0.0448$, and
%$\Omega_\nu=0$.
%Each galaxy has its cosmological redshift, the $x$-, $y$-, $z$-components of its peculiar velocity from which
%the radial peculiar velocity is determined, and its
%star formation rate and stellar mass from which the
%supernova rate of each host is determined using 
%\citet{2012ApJ...755...61S}.   This host-based rate
%underestimates the expected discovery based on the volumetric
%rate of \citet{2010ApJ...713.1026D}.  The volumetric rate is more robust, so  we boost the SN production
%by a factor 2.4 to get the 10-year volumetric rate expectation. The total number of simulated SNe 
%is greater than the total number of possibly useful supernova discoveries, since most fields cannot be continuously monitored throughout the year
%from most ground-based observatories.
%
%
%Each supernova is assigned a magnitude based on the distance modulus of its cosmological
%redshift plus a random term drawn from a Normal distribution $\sigma_M$, which for simplicity is the same for all supernovae and captures both
%intrinsic magnitude dispersion and measurement uncertainty.  No other corrections are applied to the observed magnitude.  Dipole effects of the heliocentric
%motion with respect to the CMB and of the galaxies with respect to the CMB  are ignored: referring to
%\citet{2011ApJ...741...67D}, the effects in Eq.~18 are not included.  Including these effects are straightforward but is computationally expensive
%and does not affect our conclusions.
%
%We consider a maximum redshift of $z=0.2$ because over ten years the noise is not sample-variance dominated, meaning that the differential improvement
%in the precision of
%$f\sigma_8$ due to a lower-redshift supernova is greater than that due to a high-redshift supernova.  Following all LSST SNe~Ia below this redshift 
%would likely saturate available follow-up resources.
%
%
%\section{Analysis}
%The analysis is  almost identical to that of \citet{2015JCAP...12..033H, 2017JCAP...05..015H}, except as noted later.  
%The peculiar magnitude is given by 
%\begin{equation}
%\delta_m=(m - M) - \mu(z;H_0, \Omega_M, \Omega_\Lambda).
%\label{data:eqn}
%\end{equation}  We introduce one free parameter 
%$\mathcal{M} = M - 5\log{H_0} + \text{const}$ and do not consider the other cosmological parameters, whose effects are relatively weak at low-redshift.
%
%
%The peculiar magnitude correlation
%function $\xi_{\delta m \delta m}$ \citep{2011ApJ...741...67D,2015JCAP...12..033H} expected from General Relativity acting on the CMB
%matter density power spectrum is calculated using CAMB \citep{Lewis:2002ah}  assuming the same
%cosmological parameters used for the cosmoDC2 catalog.  Our model for the data covariance
%is
%\begin{equation}
%C_{ij} = A\xi_{\delta m \delta m}(\mathbf{r_i},\mathbf{r_j}) + \frac{\sigma_M^2}{N_i} \delta_{ij} + \sigma^2_{NL}(z_i;\sigma_{v})\delta_{ij},
%\label{cov:eqn}
%\end{equation}
%where $N_i$ is the number of SNe~Ia in galaxy $i$ and $\sigma_M$ is the intrinsic SN magnitude dispersion.
%The model is used to test General Relativity through deviations from $A=1$ but is not sensitive to deviations
%from the linearized GR expectation of the shape of the power spectrum 
%$\xi$.
%Nevertheless, for
%convenience we say that our model gives $f\sigma_8 = A (f\sigma_8)^{GR}$, where $(f\sigma_8)^{GR}$ is the expectation
%from General Relativity. 
%The final term includes extra magnitude dispersion produced by non-linear effects on velocity, parameterized with $\sigma_{v}$, such that
%\begin{equation}
%\sigma_{NL} = \frac{5}{\ln{10}} \frac{1+\bar{z}}{\bar{z}} \sigma_v,
%\end{equation}
%where $\bar{z}$ is the cosmological redshift of the host galaxy.
%We note that the CAMB calculation of $\xi$ includes non-linear corrections, which actually may be a bad thing for peculiar velocities \citep{2015MNRAS.454.3920H}. 
%Such subtleties are unimportant as \citet{2015JCAP...12..033H} find that their results are insensitive to whether non-linear corrections are or are not included.
%
%For all analyses we remove supernovae below observed $z_{min}=0.01$ in order to
%reduce errors made in the first-order transformation between velocity and
%magnitude.  This cut removes a  small volume relative to the full survey.
%
%The posterior of the parameters is sampled using emcee.  The likelihood is the Normal distribution with the data in Eq.~\ref{data:eqn} and
%covariance matrix in Eq.~\ref{cov:eqn}.
%The priors are positive and flat for $A$, flat for $M$, positive Cauchy(0.08, 0.5) for $\sigma_M$, and  positive Cauchy(0, 600 km~s$^{-1}$) for $\sigma_{v}$.
%
%Direct tests of non-GR models for which Poisson's equation or the linear approximation
%does not apply requires a model-dependent $\xi$.  We defer the analysis of such higher-fidelity models to later work. 
%
%\section{Results of Subsets}
%
%The  ston in $f\sigma_8$ is calculated for different subsamples of the simulated data.
%For each subsample  the mean  and standard deviation, $(\bar{A}, \sigma_A$) of the amplitude parameter is
%determined.  The ston in $f\sigma_8$ for the subsample is thus  $\bar{A} \sigma^{-1}_A$, and that for a 18,000 sq.~deg. survey  is approximated by
%$\sqrt{18000/760}\bar{A}\sigma^{-1}_A$.  The effective constraining power of a single supernova in the subset is chracterized by 
% $\bar{A} \sigma^{-1}_A N_{gal}^{-0.5}$.
%
%Different parameters describing the input simulated data define the subsamples, 
%which are used to explore the range of possible survey strategies.   The varied parameters are the redshift range, $z_{max}$, intrinsic
%magnitude dispersion, $\sigma_M$, and SN number density, $n = \phi n_{10}$, where $ n_{10}$ is the  10-year total number density.  In general a given point
%of sky is not monitored over the full year.  
%The fit posterior for one representative subsample, with input  $z_{max}=0.2$, $\sigma_M=0.08$, $\phi=0.65$, is shown in Figure~\ref{zmax:fig}.
%For all subsamples, there is no evidence for non-convergence of the MCMC chains.  The parameter $\sigma_v$, not included in  \citet{2015JCAP...12..033H, 2017JCAP...05..015H},
%is found to be consistent with zero, poorly constrained, and very weakly correlated with $A$.  Table~\ref{tab:subsets} gives the results for the different subsamples.
%
%\begin{figure}
%\centering
%\includegraphics[width=0.5\textwidth]{../outcosmo/pvlist.0.08.1234.0.65.0.2.pkl.png}
%\caption{Confidence regions for the model parameters.  $A=(f\sigma_8)/(f\sigma_8)_{GR}$, $\mathcal{M}$ is the supernova
%absolute magnitude, $\sigma_M$ is the magnitude dispersion, and $\sigma_v$ is the non-linear contribution
%to peculiar velocity.  This result is for the subset $z_{max}=0.2$, $\sigma_M=0.08$, $\phi=0.65$.  The dotted lines represent the inputs of the supernova
%part of the simulation.
%\label{zmax:fig}}
%\end{figure}
%
%\begin{table}
%   \centering
%   %\topcaption{Table captions are better up top} % requires the topcapt package
%   \begin{tabular}{|ccc|cc|} % Column formatting, @{} suppresses leading/trailing space
%   \hline
%$z_{max}$ & $\phi$ & $\sigma_{M}$  &  STON $f\sigma_8$ & STON per SN \\
%\hline
%0.07 & 0.65 & 0.08  & 25.795 &   0.27 \\
%0.07 & 1.00 & 0.08  & 31.163 &   0.27 \\
%0.07 & 1.00 & 0.10  & 25.421 &   0.22 \\
%0.07 & 1.00 & 0.12  & 22.173 &   0.19 \\
%0.07 & 1.00 & 0.15  & 18.831 &   0.16 \\
%0.10 & 0.30 & 0.08  & 23.784 &   0.24 \\
%0.10 & 0.50 & 0.08  & 23.501 &   0.18 \\
%0.10 & 0.65 & 0.08  & 24.994 &   0.17 \\
%0.10 & 0.65 & 0.10  & 20.866 &   0.15 \\
%0.10 & 0.65 & 0.12  & 18.778 &   0.13 \\
%0.10 & 1.00 & 0.08  & 31.704 &   0.18 \\
%0.10 & 1.00 & 0.10  & 25.687 &   0.15 \\
%0.10 & 1.00 & 0.12  & 22.982 &   0.13 \\
%0.15 & 0.30 & 0.08  & 24.799 &   0.15 \\
%0.15 & 0.30 & 0.12  & 19.527 &   0.12 \\
%0.15 & 0.50 & 0.08  & 29.550 &   0.14 \\
%0.15 & 0.65 & 0.08  & 33.921 &   0.14 \\
%0.15 & 0.65 & 0.10  & 28.655 &   0.12 \\
%0.15 & 0.65 & 0.12  & 24.111 &   0.10 \\
%0.15 & 1.00 & 0.08  & 40.158 &   0.13 \\
%0.20 & 0.20 & 0.08  & 26.487 &   0.12 \\
%0.20 & 0.30 & 0.08  & 29.828 &   0.11 \\
%0.20 & 0.30 & 0.10  & 24.019 &   0.09 \\
%0.20 & 0.30 & 0.12  & 19.181 &   0.07 \\
%0.20 & 0.50 & 0.08  & 36.680 &   0.11 \\
%0.20 & 0.65 & 0.08  & 41.591 &   0.11 \\

>>>>

\documentclass{aastex62}   	% use "amsart" instead of "article" for AMSLaTeX format
\usepackage{graphicx}				% Use pdf, png, jpg, or epsÂ§ with pdflatex; use eps in DVI mode
								% TeX will automatically convert eps --> pdf in pdflatex		
\usepackage{amssymb,amsmath}
\usepackage{natbib}
\usepackage{grffile}
\usepackage{textcase}
\usepackage{hyperref}
%SetFonts

@@

The precision in  $f\sigma_8$ derived from the baseline WFD ten-year SN~Ia discoveries has been projected by \citet{2017ApJ...847..128H},
from both their RSD and peculiar velocities.
The results are summarized in Table~\ref{tab:howlett}, where a 5\% distance (0.1~mag) uncertainty is assumed for each supernova.
Their Figure~4b) shows how well these measurements can distinguish between $\gamma=0.42, 0.55, 0.68$
\citep[$f(R)$, GR,  and DGP gravities,][]{2007APh....28..481L} gravities respectively.
SNe~Ia alone can provide a $2\%$ measurement of $f\sigma_8$ at $z<0.3$ redshifts lower than where galaxy and cluster (through the
kinematic S-Z effect)
RSD measurements are sensitive.  A joint galaxy RSD and SN peculiar velocity will be even more powerful.
SNe~Ia peculiar velocities will measure $f\sigma_8$ at $z<0.3$  as well as galaxy surveys will at $z \sim 0.6$ using RSD. 
This can be compared to the current 15\% uncertainty from 6dF \citep{2017MNRAS.471..839A}, 3\% uncertainty projected
for the combined TAIPAN and WALLABY+WNSHS surveys \citep{2017MNRAS.464.2517H} for low-redshift measurements of $f\sigma_8$.
At larger $z>0.3$ redshifts DESI projects 10\% accuracies.

@@

\section{Designing a Supernova Survey to Measure $\MakeLowercase{f}\sigma_8$}
\subsection{Survey Parameters that Determine Scientific Performance}

The expected peculiar velocity signal  is described by its power spectrum $P_{vv}$.
The range $k$'s  used in an analysis depends on the depth of the survey $k_{min} \approx (2 z_{max})^{-1}$, and the smallest
scales where the theoretical predictions are reliable, say $k_{max} \approx 0.1$~$h$\,Mpc$^{-1}$.
The top panel of Figure~\ref{detC:fig} shows the volume-weighted power spectrum $k^3 P_{vv}$
at $z=0$ assuming General Relativity and linear theory.  The panel shows available $k$ ranges for several surveys with different
maximum redshift: the $0.02 <k< 0.1$~$h$\,Mpc$^{-1}$ modes contribute the bulk of the peculiar velocity signal.
The sensitivity to gravity primarily resides in the normalization of the power spectrum, as  $P_{vv} \propto (f\sigma_8)^2$.
Non-GR models can also predict differences in the shape and the redshift-dependence of GR prediction $P_{vv}$  in Figure~\ref{power:fig}.

\begin{figure}
\centering
\includegraphics[width=0.5\textwidth]{src/power.png}
%\includegraphics[width=0.5\textwidth]{../outcosmo/zmax_.png}
\caption{ 
\label{power:fig}}
\end{figure}

\citet{2017ApJ...847..128H} project how complementary galaxy and peculiar velocity surveys in a common volume can measure $f\sigma_8$. 
They perform their analysis in Fourier space, 
where the Fisher information matrix of parameters $\lambda_i$ from random Gaussian field with mean zero and covariance $C$ is
\begin{equation}
F_{ij} = \frac{V}{2}\int \frac{d^3k}{(2\pi)^3} \text{Tr}\left[ C^{-1} \frac{\partial C}{\partial \lambda_i} C^{-1}
\frac{\partial C}{\partial \lambda_j} \right]
\label{fisher:eqn}
\end{equation}
and the covariance is 
\begin{equation}
C(k,\mu)  =
  \begin{bmatrix}
   P_{gg}(k,\mu) + \frac{1}{n_g} &
   P_{vg}(k,\mu)  \\
   P_{vg}(k,\mu)  &
  P_{vv}(k,\mu) + \frac{\sigma^2}{n}
   \end{bmatrix}.
\label{cov:eq}
\end{equation}
The covariance depends on $P_{vv}$, the galaxy power spectrum $P_{gg}\propto (b\sigma_8 + f\sigma_8\mu^2)^2$, and the galaxy-velocity cross-correlation $P_{vg}
\propto  (b\sigma_8 + f\sigma_8\mu^2)f\sigma_8$, where $b$ is the bias parameter.
The galaxy number density is $n_g$, the number density of velocity probes is $n$, and the probe velocity dispersion
is $\sigma$. 
 The Fisher matrix in  Eq.~\ref{fisher:eqn} is evaluated by \citet{2017MNRAS.464.2517H}  to generate the $f \sigma_8$ uncertainties in Table~\ref{tab:howlett}.

The integrand
\begin{equation}
I = \text{Tr}\left[ C^{-1} \frac{\partial C}{\partial (f\sigma_8)} C^{-1}
\frac{\partial C}{\partial  (f\sigma_8)} \right]
\end{equation}
in Eq.~\ref{fisher:eqn}  represents how much a differential Fourier
volume element with wavenumber $k$ and line of sight $\mu$ contributes to $F_{f\sigma_{8}\,f\sigma_{8}}$;
it is therefore  useful in evaluating the design of peculiar velocity surveys. 
We evaluate these surveys relative to the $I_{max}$ of a sample-variance limited
peculiar velocity survey with\ $\sigma/n =0$.
In the following we consider as given a  TAIPAN-like
galaxy survey with $b=1.2$ and number density $n_g=10^{-3}h^3$\,Mpc$^{-3}$ that overlaps with the peculiar velocity surveys.

First, we show the benefit of having the  galaxy and velocity surveys in a common volume by considering the degradation that occurs when 
they are treated as being
independent.  For this exercise the velocity survey is taken to be sample-variance limited.
The resulting $I/I_{max}$  is shown as the black dashed curves in the bottom two panels of Figure~\ref{power:fig},
and is found to lie between $0.1-0.2$.
Taking advantage of the fact that the same dark-matter fluctuations produce the galaxies and velocities 
in the common volume decreases
the variance in $f\sigma_8$ by factors of 5 to 10.

Realistic surveys have shot noise. We consider two- and ten-year
supernova densities $n$  based on SN~Ia rates 
with 0.65 sample efficiency, and velocity dispersions $\sigma \approx (\frac{5}{\ln{10}} \frac{1+z}{z})^{-1} \sigma_M$
assuming a supernova magnitude dispersion of $\sigma_M=0.08$ and inserting values of $z=0.05,0.1, 0.2$.
The $I/I_{max}$ of these survey scenarios are shown in the bottom two panels of Figure~\ref{power:fig}.
Generally, $k<0.01$ modes have $I/I_{max}\sim 1$ meaning that they are sample-noise limited.
Individual $k>0.01$ modes have  $I/I_{max}<1$ showing that shot noise
is significant.  Recall that collectively,  larger $k$ modes are the important contributors to the peculiar-velocity signal;
basically all  supernovae found in  the next decade contribute to improving the statistical measurement of $f\sigma_8$..
The figure also shows how the statistical gain from additional supernovae diminish as the survey duration increases, and when they are at lower-redshift  (by virtue of their lower intrinsic velocity dispersion).

@@

\begin{itemize}
\item Solid Angle $\Omega$:
The variance in $f\sigma_8$ is inversely proportional to solid angle $\Omega$.  The baseline
LSST survey covers 18,000 sq.~deg, at $-75 \lesssim \delta \lesssim15$ avoiding the Galactic plane.  Larger solid-angle coverage for $z<0.3$ SN~Ia searches, beyond
the LSST baseline, benefits peculiar-velocity science.  Complementary northern-hemisphere surveys and LSST-expanded or independent
coverage of the southern equatorial pole, could significantly increase
the sky coverage with a corresponding decrease in the variance of $f\sigma_8$.
\item Redshift Depth $z_{max}$:
For the number densities generated by LSST and the SN~Ia intrinsic magnitude dispersion, the noise is not sample-variance
limited.  We have run simulations of LSST SN surveys
and calculated the signal-to-noise (STON) of $f\sigma_8$ from the peculiar-velocity correlations. These simulations show, as seen in the left plot
of  Figure~\ref{notshot:fig}, that an increase in the number density of tracers leads to a significant improvement in STON.
It is also worth noting that sample variance is not negligible, as seen by the non-linear dependence of the  STON on $\sigma_M^2/n$. 
%
%\begin{figure}
%\centering
%\includegraphics[width=0.48\textwidth]{../outcosmo/notshot.png}
%\includegraphics[width=0.48\textwidth]{../outcosmo/perSN.png}
%\caption{Left: STON of $f\sigma_8$ from simulated LSST SN peculiar-velocities only (no RSD).  There is a small level of scatter in the plots idue to the random
%realizations made in simulating the surveys.
%For several limiting redshift depths, the STON is plotted as a
%function of $\sigma_M^2/n$, normalized to intrinsic magnitude dispersion $\sigma_M=0.1$~mag and 10-year LSST number densities.
%Right: The effective contribution to the STON per supernova, quantified as STON divided by the root of the number of tracers.
%\label{notshot:fig}}
%\end{figure}

@@

A low-redshift supernova survey is insensitive to two important uncertainties
associated with the SN~Ia Hubble diagram. The measurement is relatively insensitive to absolute color calibration, due to the limited
range of {\it observer} wavelengths used to standardize restframe distance moduli.
Depending on which cameras are used to derive distances, instrumental (rather than absolute)
calibration may be sufficient.  Progenitor population  is not expected to evolve significantly in the $0.01<z<0.2$ Universe.

@@

\appendix
\section{Reanalysis in Configuration Space}

This Appendix describes a new analysis used to project precisions on $f\sigma_8$ using peculiar velocities derived from SNe~Ia.
This analysis may or may not be useful for the Decadal Survey White Paper, but will be helpful for LSST planning.
While there have been a number of articles on the subject,
our analysis brings a higher level of fidelity than sought previously.  We simulate SNe~Ia hosted by galaxies in a mock galaxy
catalog. The numbers of SNe are sufficiently small to allow fast evaluations of the likelihood, which enable the determination of parameter
posteriors using MCMC on reasonable computing timescales.   We can use our machinery to 
compare different survey parameters, such as redshift depth, total numbers of supernovae, non-uniform spatial density,
solid angle/survey geometry, and SN~Ia intrinsic magnitude dispersion.
Adding the cross-correlation between galaxy-count and peculiar-velocity surveys is active work, so we can
quantify the suppression of sample variance achieved when considering matter-densities and velocities 
within the same volume \citep{2007PhRvL..99h1301G}.

\section{Simulated Data}
The cosmoDC2 (v1.0) covering 706 sq.~deg is used\footnote{We previously used
Buzzard (v1.6) galaxy catalog is used, but it was found to give peculiar velocities inconsistent with general relativity.  As we will see,
cosmoDC2 (v1.0) is consistent with GR. }. 
The catalog is based on a Flat $\Lambda$CDM model with $H_0=71$~km~s$^{-1}$,  $\Omega_M=0.265$, $\Omega_B=0.0448$, and
$\Omega_\nu=0$.
Each galaxy has its cosmological redshift, the $x$-, $y$-, $z$-components of its peculiar velocity from which
the radial peculiar velocity is determined, and its
star formation rate and stellar mass from which the
supernova rate of each host is determined using 
\citet{2012ApJ...755...61S}.   This host-based rate
underestimates the expected discovery based on the volumetric
rate of \citet{2010ApJ...713.1026D}.  The volumetric rate is more robust, so  we boost the SN production
by a factor 2.4 to get the 10-year volumetric rate expectation. The total number of simulated SNe 
is greater than the total number of possibly useful supernova discoveries, since most fields cannot be continuously monitored throughout the year
from most ground-based observatories.

Each supernova is assigned a magnitude based on the distance modulus of its cosmological
redshift plus a random term drawn from a Normal distribution $\sigma_M$, which for simplicity is the same for all supernovae and captures both
intrinsic magnitude dispersion and measurement uncertainty.  No other corrections are applied to the observed magnitude.  Dipole effects of the heliocentric
motion with respect to the CMB and of the galaxies with respect to the CMB  are ignored: referring to
\citet{2011ApJ...741...67D}, the effects in Eq.~18 are not included.  Including these effects are straightforward but is computationally expensive
and does not affect our conclusions.

We consider a maximum redshift of $z=0.2$ because over ten years the noise is not sample-variance dominated, meaning that the differential improvement
in the precision of
$f\sigma_8$ due to a lower-redshift supernova is greater than that due to a high-redshift supernova.  Following all LSST SNe~Ia below this redshift 
would likely saturate available follow-up resources.

\section{Analysis}
The analysis is  almost identical to that of \citet{2015JCAP...12..033H, 2017JCAP...05..015H}, except as noted later.  
The peculiar magnitude is given by 
\begin{equation}
\delta_m=(m - M) - \mu(z;H_0, \Omega_M, \Omega_\Lambda).
\label{data:eqn}
\end{equation}  We introduce one free parameter 
$\mathcal{M} = M - 5\log{H_0} + \text{const}$ and do not consider the other cosmological parameters, whose effects are relatively weak at low-redshift.

The peculiar magnitude correlation
function $\xi_{\delta m \delta m}$ \citep{2011ApJ...741...67D,2015JCAP...12..033H} expected from General Relativity acting on the CMB
matter density power spectrum is calculated using CAMB \citep{Lewis:2002ah}  assuming the same
cosmological parameters used for the cosmoDC2 catalog.  Our model for the data covariance
is
\begin{equation}
C_{ij} = A\xi_{\delta m \delta m}(\mathbf{r_i},\mathbf{r_j}) + \frac{\sigma_M^2}{N_i} \delta_{ij} + \sigma^2_{NL}(z_i;\sigma_{v})\delta_{ij},
\label{cov:eqn}
\end{equation}
where $N_i$ is the number of SNe~Ia in galaxy $i$ and $\sigma_M$ is the intrinsic SN magnitude dispersion.
The model is used to test General Relativity through deviations from $A=1$ but is not sensitive to deviations
from the linearized GR expectation of the shape of the power spectrum 
$\xi$.
Nevertheless, for
convenience we say that our model gives $f\sigma_8 = A (f\sigma_8)^{GR}$, where $(f\sigma_8)^{GR}$ is the expectation
from General Relativity. 
The final term includes extra magnitude dispersion produced by non-linear effects on velocity, parameterized with $\sigma_{v}$, such that
\begin{equation}
\sigma_{NL} = \frac{5}{\ln{10}} \frac{1+\bar{z}}{\bar{z}} \sigma_v,
\end{equation}
where $\bar{z}$ is the cosmological redshift of the host galaxy.
We note that the CAMB calculation of $\xi$ includes non-linear corrections, which actually may be a bad thing for peculiar velocities \citep{2015MNRAS.454.3920H}. 
Such subtleties are unimportant as \citet{2015JCAP...12..033H} find that their results are insensitive to whether non-linear corrections are or are not included.

For all analyses we remove supernovae below observed $z_{min}=0.01$ in order to
reduce errors made in the first-order transformation between velocity and
magnitude.  This cut removes a  small volume relative to the full survey.

The posterior of the parameters is sampled using emcee.  The likelihood is the Normal distribution with the data in Eq.~\ref{data:eqn} and
covariance matrix in Eq.~\ref{cov:eqn}.
The priors are positive and flat for $A$, flat for $M$, positive Cauchy(0.08, 0.5) for $\sigma_M$, and  positive Cauchy(0, 600 km~s$^{-1}$) for $\sigma_{v}$.

Direct tests of non-GR models for which Poisson's equation or the linear approximation
does not apply requires a model-dependent $\xi$.  We defer the analysis of such higher-fidelity models to later work.

\section{Results of Subsets}

The  ston in $f\sigma_8$ is calculated for different subsamples of the simulated data.
For each subsample  the mean  and standard deviation, $(\bar{A}, \sigma_A$) of the amplitude parameter is
determined.  The ston in $f\sigma_8$ for the subsample is thus  $\bar{A} \sigma^{-1}_A$, and that for a 18,000 sq.~deg. survey  is approximated by
$\sqrt{18000/760}\bar{A}\sigma^{-1}_A$.  The effective constraining power of a single supernova in the subset is chracterized by 
 $\bar{A} \sigma^{-1}_A N_{gal}^{-0.5}$.

Different parameters describing the input simulated data define the subsamples, 
which are used to explore the range of possible survey strategies.   The varied parameters are the redshift range, $z_{max}$, intrinsic
magnitude dispersion, $\sigma_M$, and SN number density, $n = \phi n_{10}$, where $ n_{10}$ is the  10-year total number density.  In general a given point
of sky is not monitored over the full year.  
The fit posterior for one representative subsample, with input  $z_{max}=0.2$, $\sigma_M=0.08$, $\phi=0.65$, is shown in Figure~\ref{zmax:fig}.
For all subsamples, there is no evidence for non-convergence of the MCMC chains.  The parameter $\sigma_v$, not included in  \citet{2015JCAP...12..033H, 2017JCAP...05..015H},
is found to be consistent with zero, poorly constrained, and very weakly correlated with $A$.  Table~\ref{tab:subsets} gives the results for the different subsamples.

\begin{figure}
\centering
\includegraphics[width=0.5\textwidth]{../outcosmo/pvlist.0.08.1234.0.65.0.2.pkl.png}
\caption{Confidence regions for the model parameters.  $A=(f\sigma_8)/(f\sigma_8)_{GR}$, $\mathcal{M}$ is the supernova
absolute magnitude, $\sigma_M$ is the magnitude dispersion, and $\sigma_v$ is the non-linear contribution
to peculiar velocity.  This result is for the subset $z_{max}=0.2$, $\sigma_M=0.08$, $\phi=0.65$.  The dotted lines represent the inputs of the supernova
part of the simulation.
\label{zmax:fig}}
\end{figure}
